### 2018-03-05 abstract.tex
compiles

A critically important, but surprisingly neglected in the software testing literature, aspect of system reliability is system predictability.  Many software systems are implemented using mechanisms (unsafe languages, mulitple threads, complex caching, stochastic algorithms) that can introduce behavioral non-determinism.  Users of software systems, including especially other software systems using library calls in a single-threaded context, however, expect that systems will behave deterministically from the perspective of having predictable results from the same sequential series of operations.  In this paper we show that lightweight techniques, requiring (almost) no effort on the part of a developer, can extend an existing test generation system to allow checking of determinism properties on tested code.  We introduce two notions of determinism to check (visible value trace determinism and failure determinism), inspired by real faults, and tailor these notions to the practical, automatic, checking of Python library code.