### 2018-03-05 relatedwork.tex
small additions

\section{Related Work}

Shi et. al \cite{DetermImp} examined what might be seen as a related, though in a sense the opposite, problem:  they detect code that assumes a deterministic implementation of a non-deterministic specification.  E.g., they detect instances when  iteration through items of a set is explicitly not guaranteed to provide any particular order of items, but code depends on the order produced by a given implementation.