### 2018-03-05 abstract.tex
better abstract

A critically important, but surprisingly neglected in the software testing literature, aspect of system reliability is system predictability.  Many software systems are implemented using mechanisms (unsafe languages, mulitple threads, complex caching, stochastic algorithms, environmental dependencies) that can introduce behavioral nondeterminism.  Users of software systems, especially other software systems using library calls in a single-threaded context, often expect that systems will behave deterministically in the sense that they have predictable results from the same sequential series of operations.  Equally importantly, even when it does not impact correctness or user experience, nondeterminism must be understood and managed for effective automated testing.  Nondeterministic behavior that is not either controlled or accounted for can result in flaky tests as well as causing problems for test reduction, differential testing, and automated regression test generation.  We show that lightweight techniques, requiring (almost) no effort on the part of a developer, can be used to extend an existing test generation system to allow checking for problematic nondeterminism.  We introduce three specific lightweight nondeterinism properties (visible value trace and final-value nondeterminism and failure nondeterminism), inspired by real faults, and tailor these notions to the practical, automatic, checking of Python library code.