### 2018-03-06 relatedwork.tex
more related work

\section{Related Work}

Shi et. al \cite{DetermImp} examined what might be seen as a related, though in a sense the opposite, problem:  they detect code that assumes a deterministic implementation of a non-deterministic specification.  E.g., they detect instances when  iteration through items of a set is explicitly not guaranteed to provide any particular order of items, but code depends on the order produced by a given implementation.

Other efforts \cite{ParallelDeterministic} have aimed to avoid nondeterminism in parallel implementations, by design, indicating the importance of avoiding nondeterministic behavior, even at the expense of adopting novel programming models, where the goal is not simply (as in, say, Rust) to avoid classic concurrency errors, but to enforce genuinely deterministic behavior.