### 2018-03-24 introduction.tex
more text

\section{Introduction}

For ourselves, we might prefer to think (and act as if) we had free
will; however, we generally prefer our software systems to be as
constrained in their actions as possible:  in other words, we wish
them to be largely deterministic, from our perspective.

WHY DETERMINISM:  Bugs, flaky tests, pain debugging.

\subsection{What is Determinism?}

A system is deterministic if, given its complete state at a point in
time, it is possible, in principle, to predict its future behavior
perfectly.  We say ``in principle'' because in the real world,
prediction may be possible but hopelessly impractical.  We write
complex software systems in many cases because we cannot predict their
behavior (if we could perform the calculations in advance, ourselves,
we would just do so).  As a consequence, in software, rather than
defining determinism in terms of prediction, we usually therefore
simply say that a system is deterministic if, given the same state and
inputs, it always produces the same outputs\footnote{Of course, this
  shift of definition is often adopted in arguments for physical or
  human behavioral determinism, due to the even greater difficulties
  of prediction in those arenas.}.  The change of
definition, we will see, is central to our approach, and comes with a
number of limitations.

Technically, many ``nondeterministic'' systems are not
nondeterministic in a strong sense, at all.  Given the complete state
of the system (which includes the entire state of the underlying
hardware, the operating system, storage devices, etc.), ignoring
quantum effects, and treating outside interventions such as network
traffic, human activity at an input device, etc. properly as inputs,
most software \emph{is} completely predictable.  What we actually
mean, usually, is that, given a certain limited abstraction of state
and of inputs, observable behavior is repeatable.  This abstraction,
for most software, is not expected to include many elements outside of
the software system itself.

Unexpected nondeterminism usually arises when there is an element of
state or input that \emph{is} critical to the produced behavior, but
the programmer has not anticipated.  E.g., it is beleived that the
behavior of a thread scheduler will not matter, but a race condition
in the code means that it does matter, after all.

\subsection{The High Cost of Unexpected Nondeterminism}

Unexpected nondeterminism is, unfortunately, usually only discovered
in a context that makes it very hard to debug.   The most common such
contexts are occasional rare failures of a system in deployment, and
flaky tests.