### 2019-09-04 experiments.tex
edit exp

\section{Experimental Evaluation}

In order to evaluate our approach and implementation, we applied
nondeterminism detection to real Python code, including some widely
used systems ({\tt redis-py} and {\tt pyfakefs} are well-maintained,
widely-used libraries).  The primary
points we wanted to explore were (1) whether our approach was able to reliably detect actual
nondeterminism, (2) whether the overhead of nondeterminism
detection for real systems was acceptable, and (3) the performance of
delta-debugging for real nondeterministic predicates.  We chose two of
the subjects, {\tt redis-py} and {\tt pyfakefs}, based on the fact that they were previously-existing large
examples of TSTL harnesses, and therefore serve to show how easily our
approach can be integrated into an existing large-scale test
generation effort.  The other two subjects were chosen, and harnesses
written, specifically to demonstrate specific, important sources of
nondeterminism in Python code: concurrency and hash salting.

Table \ref{tab:overviewexp} shows a summary of our experimental
results.  In all cases, our approach was able to detect interesting
nondeterminism.  The overhead on testing imposed by nondeterminism
checks varied considerably, but in the worst case did not double the
average cost of each action, and thus never halved the amount of
testing performed.  Test reduction always reduced the size of
generated tests by 90\% or more, thus showing that effective reduction
has a large payoff.

\begin{table}
\centering
\begin{tabular}{l|l|r|r|l}
Subject & Detection & Overhead & Reduction & Reduction time \\
\hline
Parallel Sort & Yes (see below) & $\sim$40\% & & $\sim$2 minutes\\
{\tt redis-py} & Yes (see below) & $\sim$20\% & $\sim$95\% & $\sim$10 minutes\\
{\tt datarray} & Yes (see below) & $\sim$93\% & $\sim$90\% & $\sim$92
  seconds\\
{\tt pyfakefs} & Yes (see below) & $\sim$8\% & $\sim$99\% & $\sim$1 second  \\
\end{tabular}
\caption{Overview of experimental results}
\label{tab:overviewexp}
\end{table}

\subsection{Parallel Sorting}

Our first subject is a simple implementation of a parallel merge
sort\footnote{\url{https://gist.github.com/stephenmcd/39ded69946155930c347}}.  The
merge sort and harness can be found in the TSTL GitHub repository
\cite{tstl}, under {\tt examples/parallelsorts}.  When a sorting
algorithm is implemented using concurrency, a key question to ask is
whether the ordering of elements that are equal under the comparison
operator is consistent across sorts of the same original sequence.
This is related to the question of whether a sort is a stable sort
(preserving the original relative order of equal-by-comparison elements),
but imposes a less strict requirement on the sort; a concurrent sort
might not be stable, but might still produce the same (non-stable)
ordering from the same provided sequence, every time.  This is a
particular instance of a common expectation noted above:  while a parallel
implementation of an algorithm will likely have internal
nondeterminism (due to scheduling of work), it is often expected to
behave deterministically, from the point of view of a caller.

This experimental subject serves two purposes:  first, to show that the
nondeterminism detection features of TSTL can be used to, with
acceptable overhead, show that a parallel sort implementation provides
a consistent, deterministic ordering, and second to show that the
mechanism can also detect the nondeterminism in a parallel sort that
does exhibit nondeterministic behavior.  For the second purpose, we
implemented a very simple swap-based parallel sorting algorithm that
repeatedly has multiple threads scan through a sequence and swap
out-of-order neighbors, until the sequence is sorted (a kind of
parallel bubble-sort with a simpler termination criteria).

\begin{figure}
{\scriptsize
\begin{code}
@import mergesort
@import swapsort
\vspace{0.1in}
pool: <val> 10
pool: <data> 10
pool: <sorted> 10
\vspace{0.1in}
<val> := <[-10..10]>
\vspace{0.1in}
<data> := []
<data>.append((<val>,<val>))
len(<data,2>) < len(<data,1>) -> <data>.extend(<data>)
\vspace{0.1in}
<sorted> := mergesort.merge\_sort\_parallel(<data>)
<sorted> := swapsort.swap\_sort\_parallel(<data>)
<sorted>
\vspace{0.1in}
property: swapsort.is\_sorted(<sorted>)
\end{code}
}
\caption {Complete TSTL harness for checking two parallel sorting
  algorithms for nondeterminism.}
\label{fig:parallelcode}
\end{figure}

Figure \ref{fig:parallelcode} shows the complete TSTL harness for
checking the parallel sorts.  This harness produces sequences of
integer pairs, where only the first integer is used in the comparison
operator of the sorting algorithms, and checks that the results of
{\tt sort\_parallel} calls are in fact sorted, using the {\tt
  swapsort}'s function for checking if a sequence is sorted.

Running TSTL's random tester without checking for determinism, we can
see that both sorting algorithms always produce sorted output.  If we run with {\tt --checkDeterminism --determinismDelay 0}, however, a
test case showing the {\tt swapsort} behaving nondeterministically is
almost instantly produced (it takes less than 3 seconds).  If we comment out the call to {\tt
  swap\_sort\_parallel} and only check the parallel merge sort, however, TSTL
finds no nondeterministic behavior.  Running tests of the merge sort
with nondeterminism detection turned on results in a slowdown of approximately
40\%.

Reducing a lengthy test showing nondeterminism of the swapsort to a
short test case using delta-debugging usually takes less than two
minutes.  If the number of attempts to produce nondeterminism is
increased in the reducer using the {\tt --determinismTries 10} option,
the resulting test will be usually be very compact, e.g.:

{\scriptsize
\begin{code}
data0 = []    
val0 = 3      
val1 = -9    
data0.append((val0,val1)) 
val2 = -1                         
data0.append((val1,val2))
data0.append((val1,val1))  
sorted0 = swapsort.swap\_sort\_parallel(data0)      
\end{code}
}

For this simple example, more sophisticated delta-debugging strategies
are not needed; we only want to see the code behaving
nondeterministically, and do not care about preserving a higher probability of nondeterminism.

\subsection {The {\tt redis-py} Library}

\begin{figure*}
\centering 
\begin{subfigure}{0.3\columnwidth}
\centering
\includegraphics[width=\columnwidth]{redisddmin}
\caption{Standard delta-debugging (38s)}
\label{fig:r1}
\end{subfigure}
\begin{subfigure}{0.3\columnwidth}
\centering
\includegraphics[width=\columnwidth]{redisforcep}
\caption{P=0.5, 100 samples (2131s)}
\label{fig:r2}
\end{subfigure}
\begin{subfigure}{0.3\columnwidth}
\centering
\includegraphics[width=\columnwidth]{redisforceprep}
\caption{P=0.5, 10 samples, 10 replications (617s)}
\label{fig:r3}
\end{subfigure}
\caption{Delta-debugging of the same {\tt redis-py} nondeterministic test}
\end{figure*}

The {\tt redis-py} \cite{redispy} module implements a widely-used Python interface
to the popular Redis in-memory database, cache, and
message broker \cite{redis}.  The TSTL {\tt redis-py} harnesses can be found in
the {\tt examples/redis} directory in the TSTL GitHub repo \cite{tstl}.

Using TSTL's harness for {\tt redis-py}, both {\tt redis-py} and Redis
itself can be tested; unfortunately, generating stand-alone
high-coverage regression tests for {\tt redis-py} and Redis using TSTL
has proven difficult, as numerous Redis commands introduce nondeterministic
behavior:  thus the resulting tests are very often \emph{flaky}.   Using {\tt tstl\_afl\_fuzz}, we can see that {\tt redis-py} has a 
stability of only 56.26\%, a clear indicator of nondeterminism. Some
of the problematic commands are obvious simply by inspection (e.g.,
{\tt randomkey, srandmember}).  And, given an understanding of Redis
semantics, it is also clear that the various
commands producing data with a limited lifetime (e.g., {\tt expire,
  pexpire}) also introduce timing-based nondeterminism.

However, that a command such as {\tt restore} takes an expiration
argument is not obvious to a non-expert, nor is the behavior of {\tt
  spop} which pops a random value from a set.  Moreover, it is
difficult to guess whether the {\tt pipe} mechanism, which allows a
large sequence of commands to be queued up in a pipe and executed all
at once, introduced the potential for nondeterminism (does it execute
sequentially before other command are handled, or is it in parallel
with further commands).  Deep experience with Redis would make these
issues clear, but tests using a library are often written by those not
intimately familiar with the semantics of every library call (otherwise they would not have
been as likely to introduce flaky tests in the first place).  This is
even more true in the case of automated test generation, where the
test engineer is often chosen for expertise in test generation tools,
not in the domain of the software under test (SUT), and is seldom the
original developer of the code.

Using the original TSTL harness, just over 20\% of all {\tt redis-py}
regression tests (of length 200) generated were flaky.  Using the
nondeterminism checker to reduce flaky tests to a minimal set allows
us to simply set up an overnight run that will, in 12 hours, produce
a set of minimal nondeterministic tests exposing the sources of
nondeterminism/flakiness.  This process is unlikely to require 12
hours, of course; our point is that it is completely automated,
including generation of a set of minimal, distinct, tests showing
different kinds of nondeterminism using test normalization~\cite{onetest}.  The actual amount of time given
may be smaller, or, for that matter, larger (leaving an important
library to self-document nondeterminism over a weekend is hardly a bad
practice).  In this case, the probability of flakiness
is not important, just the possibility, so we let the delta-debugging
create a minimal test with some non-zero (but possibly very small)
observed probability of nondeterminism.  Our 12 hour run identified
all sources of nondeterminism successfully, according to a series of
48 hour runs performed on the version of the harness we modified to
remove nondeterminism.  We determined that the mean time to find all
the sources of nondeterminism and produce minimal tests 3.2
hours, which is not trivial, but also, for this purpose, not excessive.

Figures \ref{fig:r1}-\ref{fig:r3} show delta-debugging of a typical
{\tt redis-py} nondeterministic test, originally with 287 steps.  The initial test behave
nondeterministically about half the time.  Figure \ref{fig:r1} shows
that this is a non-monotonic reduction problem, where removing steps
can either increase or decrease the probability of nondeterminism.  At
first, unmodified delta-debugging actually improves the probability of
failure, but eventually it produces a test of length 16 that only behaves
nondeterministically 24\% of the time.  The reduction takes only 38
seconds.  For the purpose of identifying commands leading to
nondeterminism, this is acceptable.  However, if we were actually
debugging a complex nondeterminism bug in Redis itself, we might want
a more reliably nondeterministic test.  Using the same parameters as
in Section \ref{sec:pubbias}, we see the same pattern.  Simply making
a predicate that ``forces'' the probability to remain high, with a
large number of samples, does not work (Figure \ref{fig:r3}) and
requires over 2000 seconds to produce a test with an even worse
probability of nondeterminism.  Using 10 samples and 10 replications,
on the other hand, actually improves the probability of
nondeterministic behavior, and
(in this case) even produces a smaller test (only 14 steps) in just
over 10 minutes.

\begin{figure*}
{\scriptsize
\begin{code}
{\bf REMOVED:}    
<key> := <r>.randomkey() 
<r>.expire(<key>,<int>)
<r>.pexpire(<key>,<int>)
\{redis.exceptions.ResponseError\} <r>.spop(<key>)
\{redis.exceptions.ResponseError\} <r>.srandmember(<key>)
\{redis.exceptions.ResponseError\} <r>.srandmember(<key>,<int>)
<pipe>.expire(<key>,<int>)
<pipe>.pexpire(<key>,<int>)
\{redis.exceptions.ResponseError\} <pipe>.spop(<key>)
\{redis.exceptions.ResponseError\} <pipe>.srandmember(<key>)
\{redis.exceptions.ResponseError\} <pipe>.srandmember(<key>,<int>)
\vspace{0.1in}
{\bf MODIFIED:}
\{redis.exceptions.ResponseError\} <r>.restore(<key>,<int>,<dump>)
{\bf $\Rightarrow$}
\{redis.exceptions.ResponseError\} <r>.restore(<key>,0,<dump>)
\vspace{0.07in}
\{redis.exceptions.ResponseError\} <pipe>.restore(<key>,<int>,<dump>)
{\bf $\Rightarrow$}
\{redis.exceptions.ResponseError\} <pipe>.restore(<key>,0,<dump>)
\end{code}
}
\caption{{\tt redis-py} calls removed or changed to avoid
  nondeterminism.}
\label{fig:flakysource}
\end{figure*}

Figure \ref{fig:flakysource} shows all calls removed or modified after
identification of sources of nondeterminism.  After making these changes, no flakiness was observed in a sample of over 2,000 length
200 tests.   Moreover, because the removals were limited to actually
observed sources of flakiness, the code coverage loss was minimal.
Mean branch coverage, for regression tests of length 200 was reduced
by less than 5 branches (a less than 1\% decrease).  Obviously, it is
impossible to test the removed calls, but the overall coverage loss is
both minimal and known, and can be made up for using specially-crafted
tests (for instance, wrapping the problematic calls in a way that does
not check the values, or adding a delay after an expiration to allow
the data to expire).  As a price to pay for the ability to produce
fast-executing high-coverage full regression tests (not just tests for
crashes and unexpected exceptions), this seems acceptable.

Moreover, turning off sources of expected nondeterminism makes
it possible to aggressively test {\tt redis-py} and Redis for
unexpected nondeterminism arising from actual bugs.  The overhead for such
determinism checks, with no delay between operations, is only
20\%, much lower than the expected cost of running each test twice.
This is because choosing the actions in a test (and determining which
actions are enabled at each step) consumes a large part of the test
generator's time; running the test again and checking equality is
relatively inexpensive.

Finally, AFL stability was 94.32\% with the harness allowing
nondeterminism, but rose to 98.32\% using the harness with
nondeterministic operations removed.

\subsubsection{Visible Value vs. Final State Nondeterminism}

We also used {\tt redis-py} to investigate the tradeoff between
overhead and ability to detect nondeterminism when using the two
proposed approaches to horizontal determinism.  We produced 300 tests
of length 100, using the known-nondeterministic version of the {\tt
  redis-py} harness.  We then used {\tt nondeterministic} and {\tt
  stepNondeterministic} calls, both with a delay of 0.005 seconds and
a single additonal execution of the test, to
check these tests for nondeterministic behavior.  Final state
nondeterminism actually detected one instance of nondeterminism that
visible value nondeterminism failed to detect (21 detections vs. 20); in general, final state
nondeterminism is less able to detect nondeterminism, but in this
instance the difference is less important than the nondeterminism of
whether a particular test will exhibit nondeterministic behavior in a
single run.  This difference is obviously not statistically
significant; for the nondeterminism in {\tt redis-py}, with these
parameters, the two approaches cannot be statistically distinguished
as to effectiveness in detecting nondeterminism.

Final state nondeterminism was also faster; visible value
nondeterminism took, on average, 0.48\% longer to check for
nondeterminism on the 300 tests, 1.199 mean seconds vs. 1.205 mean seconds.  The difference was significant, with
$p$-value $< 1.19 \times 10^{-23}$ by a paired Wilcoxon test.

These results, of course, depend on the parameters of the check for
nondeterminism.  If we increase the delay to 0.01 seconds and the
number of replays of a test used to check for nondeterminism, we
almost double the number of nondeterministic tests discovered.  Both
visible value and final state approaches find 40 nondeterministic
tests, though the overlap of tests thus detected is not quite perfect
--- each method detects two tests the other does not.  The difference
in overhead, interestingly, is very similar:  0.48\%; however, under
these parameters, visible value nondeterminism is actually cheaper on average
than final state nondeterminism, due to early termination of the
additional replays, when nondeterminism is detected (the difference is
again significant by Wilcoxon test, with $p < 1.44 \times 10^{-10}$.

In practice, visible value nondeterminism and final state
nondeterminism will often be very similar in both ability to detect
nondeterministic behavior and overhead.  However, under unusual
conditions, the behavior of the approaches can be very different.  The
additional overhead of visible value nondeterminism is usually low
because re-executing a test, not comparing state values for equality, is
the primary cost in nondeterminism detection; however, if there are a
very large number of state components, or some state components have
an expensive equality check (e.g., recursive structures with cycle
detection or depth limits), the overhead can grow, proportional to the
length of tests under consideration.  Similarly, final state
nondeterminism in a setting such as {\tt redis-py}, where divergences
in behavior tend to propagate to other state components, or at least
persist until termination of a test, detects nondeterminism quite
effectively.  However, in a setting where changes in behavior can
easily by overwritten by future test behavior, and do not causally
influence future computation, final state nondeterminism may be almost useless.

\subsection{Berkeley {\tt datarray} Inference Algorithms}

\begin{figure}[t]
{\scriptsize
\begin{code}
@import inference\_algs
@import datarray
\vspace{0.1in}
<@
def flatten\_and\_sort(v):
    return (sorted(map(flatten\_and\_sort,v),key=repr) if type(v) in [list,tuple] else
                (flatten\_and\_sort(list(v.items())) if type(v) == dict else v))

def psplit(P):
    return ([P,1.0-P])
@>
\vspace{0.1in}
pool: <P> 3
pool: <cpts> 3
pool: <evidence> 3 OPAQUE
pool: <ename> 3
pool: <event> 3
\vspace{0.1in}
<P> := 0.01 * <[0..100]>
<ename> := "E" + str(<[1..5]>)
\{Exception\} <event> := datarray.DataArray(psplit(<P>), axes = [<ename>])
\{Exception\} <ename,1>!=<ename,2> -> <event> := [datarray.DataArray([[psplit(<P>)],
   psplit(<P>)],[<ename>,<ename>])]
<cpts> := []
~<cpts>.append(<event>)
<evidence> := \{\}
~<evidence>.update([(<ename>,0)])
\vspace{0.1in}
\{Exception\} print(flatten\_and\_sort(inference\_algs.calc\_marginals\_simple(<cpts>,
   <evidence>)))
\{Exception\} print(flatten\_and\_sort(inference\_algs.calc\_marginals\_sumproduct(<cpts>,
   <evidence>)))
\{Exception\} print(flatten\_and\_sort(inference\_algs.calc\_marginals\_jtree(<cpts>,
   <evidence>)))
\end{code}
}
\caption {Complete TSTL harness for finding the hash-order bug in the datarray
  inference algorithms.}
\label{hashbug}
\end{figure}

The {\tt datarray} module \cite{datarray} is a prototype
implementation for numpy arrays with named axes to improve data
management, developed by the Berkeley Institute for Data Science.  As part of its code, it provides a set of algorithms for
inference in
Bayesian belief networks \cite{russell2016artificial}.  An earlier
version of these algorithms produced nondeterministic (and in some
cases incorrect) results due to dependence on the order of values in
an iterator over a Python dictionary, on Python versions above 3.2,
until 3.6 (see Section \ref{sec:pnondet}).

Figure \ref{hashbug} shows TSTL code for generating inputs to the {\tt
  datarray} algorithms (it can be found in the {\tt
  examples/datarray\_inference} directory in the TSTL repository \cite{tstl}).  The {\tt flatten\_and\_sort} function is needed because we care about actual differences in probability values, not simply the order of list, tuple, or dictionary items.  Running this harness using TSTL's {\tt
  --checkProcessDeterminism} flag required less than 10 seconds on
average to produce a test exhibiting process-level nondeterminism in
the {\tt calc\_marginals\_sumproduct} function (the only broken
algorithm).  Reducing this 60 step test to a minimal test of only 6 steps,
showing an extremely simple input producing the issue, 
required another 92 seconds.  Interestingly, the way that {\tt
  python-afl} is used in TSTL means that AFL stability is 100\%,
because the {\tt PYTHONHASHSEED} has already been chosen before the
fork in AFL executions.

Removing the nondeterministic call, we can see that the cost of
checking for process nondeterminism, with no delay between operations,
is high, a 93\% slowdown.  This is due to the high cost of subprocess
creation and communication.

\subsection {Vertical Determinism: {\tt pyfakefs}}

The {\tt pyfakefs} \cite{pyfakefs} module implements a fake file
system that mocks the Python file system modules, to allow Python
tests both to run faster by using an in-memory file system and to make
file system changes that would not be safe or easily performed using
real persistent storage.  Originally developed in 2006 at Google by
Mike Bland, {\tt pyfakefs} is now used in over 2,000 Python tests,
inside and outside Google \cite{pyfakefs}.

The TSTL harness for {\tt pyfakefs} has been used to detect (and
correct) over 80 faults.  The harness can be found in the TSTL GitHub repository
in the {\tt examples/pyfakefs} directory, and the bugs discovered can
be viewed at \url{https://github.com/jmcgeheeiv/pyfakefs/issues?q=label%3ATSTL}.  However, the testing largely relies on
the existence of a reference file system implementation.  One purpose
of failure nondeterminism is to make it somewhat easier to perform effective
property-based testing of
complex APIs like this even without a complete reference implementation, or in cases
where the implementations do not use the same error
codes (as is common, e.g. in NASA flight software
\cite{ICSEDiff,CFV08}).

\subsubsection{Manually Inserted Fault}

We introduced a subtle bug into {\tt pyfakefs}, where the {\tt remove}
call checks that its target is not a directory, and returns the
correct error, but still carries out the remove operation.  Using {\tt
  os.remove} to delete directories does not break any file system
invariants, but violates the Python {\tt os} specification (and,
indirectly, the usual POSIX implementation behavior where {\tt unlink}
does not work for directories).  Detecting this bug using the TSTL
{\tt pyfakefs} harness is normally impossible without using another
file system as a reference.  However, the fault was detected
immediately, even without using a reference, when we compiled the harness
with the {\tt --checkFailureDeterminism} flag.  Because vertical
reduction does not require running complete tests multiple times, and
does not affect the delta-debugging algorithm's performance, reducing
the failing test to 3 steps required less than a second.
Moreover, the overhead for the check for failure determinism in a
version of the code without the {\tt remove} error was
less than 8\%.  Detecting the fault using a reference file system 
required 17\% more testing time before detection, and took over twice as
long to reduce the failure to a
slightly longer failing test, which did not have {\tt remove} as its
final operation (since further operations are required to 
expose the bad file system state the operation introduces).  For this
hypothetical subtle bug, failure determinism checks not only make it
possible to detect the fault without a reference implementation, it
improves on detection speed and ease of debugging even compared to a
full-blown reference implementation (the MacOS file system, operating
on a RAM disk) and strong, hand-tuned, differential testing.

\subsubsection{Mutation Analysis}

We wanted to determine whether there were also simple faults that
could be detected by failure nondeterminism, but \emph{not} detected
by differential testing, and quantify the additional specification
strength provided by failure determinism checks.  We therefore used
{\tt universalmutator} \cite{RegExpMut} to produce 2,350 mutants of
the {\tt pyfakefs} core file system code.   We restricted the generation
to only mutate code covered during a 60 second run of the test
harness, with differential testing turned off.  In this
\emph{non-differential} mode, the harness will only detect a fault when it
causes an unexpected exception to be raised, or results in a timeout
due to, e.g., an infinite loop.

We first analyzed the mutants using 60 seconds of non-differential
testing \emph{without failure nondeterminism checks}, then with the
full differential harness (also without failure nondeterminism checks)
for 120 seconds; the additional 60 seconds is to make sure we accounted for the observed 17\%
extra time to detect in the manually constructed example: we want to
maximize the chance to detect a bug using the strong version of the harness.   For both comparisons, we then tested
all surviving mutants using the non-differential harness for 60
seconds, this time with failure nondeterminism checking turned on.
We used the same random seed for
all runs, so that the only differences would be specification-based.

Non-differential testing, without failure determinism checks, killed
872 mutants, a 37.1\% kill rate.  Adding a failure determinism check
allowed the testing to kill an additional 98 mutants, an 11\% improvement.
The differential harness, with an additional 60 seconds of testing
time, killed 1148 mutants, improving the kill rate to 48.8\%.  Failure
nondeterminism checking added an
additional 71 mutants to the total killed, a 6\% improvement even for
this strong differential harness with a larger test budget.  Using
failure nondeterminism was the only way to push the mutation score
above 50\%.  The kill rates are generally low
because {\tt universalmutator}, by design, includes many operations
that can produce equivalent mutants, but can also produce hard-to-kill
non-equivalent mutants not produced by other mutation tools, e.g. code
that throws away exceptions raised by a function call (see below), or code
reversing a list.

In order to further investigate the additional specification power
provided by failure nondeterminism detection, we inspected the mutants
killed using the failure determinism check but not killed by the
strong differential testing.  The largest category of mutants not
killed (22 of the 71 mutants) was what we refer to as ``exception
swallowing'' mutants, which transform a Python statement into the same
statement, but wrapped in a {\tt try} block with a catch that ignores
any raised exceptions, e.g., {\tt foo()} becomes:

\begin{code}
try: foo()
except: pass
\end{code}

\noindent It is easy to
see that such mutants may introduce faults in the handling of
errors, and thus would tend to cause failure nondeterminism.  However,
these are a minority of the mutations introducing hard-to-detect but
non-equivalent failure
nondeterminism.  Other mutation operators resulting in subtle flaws
not (at least easily) detectable by differential testing compared to a
correct file system include:  arithmetic operation changes, statement
deletions, logical operator modifications, constant replacements
(including replacement of a string with the empty string), and
introducing a break into a loop.  The variety of mutant types suggests
that no more specifically tailored strategy such as checking for
exception propagation, will work as well as introducing a notion of
failure nondeterminism.

We also checked the cost of introducing failure determinism checking
by analyzing mutants that could be killed by both methods.  Even
though in some cases the failure determinism check allows a mutant to
be detected sooner, the mean time to kill mutants was 0.006 seconds
larger with failure determinism checking, and this change was significant by Wilcoxon test
$(p < 1 \times 10^{-15})$.  While significant, this cost is probably
too small to be of any real practical importance.

\subsection{Summary of Results}

\section{Threats to Validity}

There are several primary threats to validity:  first, our empirical
results are limited to a small set of Python programs, ranging from
relatively small and simple to large and complex libraries; the
representative nature of these subjects is not clear.
Furthermore, because
no other tools implement the kind of approach taken here, based on
automated generation of value pool based unit tests for a system
(imperative property-based testing), to our knowledge, we were unable
to perform a meaningful comparison with another nondeterminism
detection tool.  Available tools, such as DeFlaker \cite{bell2018d}
(\url{http://www.deflaker.org/}), perform a completely different task (DeFlaker
is essentially a plugin for Maven regression tests, and cannot check
nondeterminism at any smaller granularity than that of a whole tests's
pass/fail result) and target the Java language.  iDFlakies
\cite{idflakies}, similarly, ``does not further classify the causes of
the flaky tests'' beyond identifying them as likely related to test
order, or not.

However, the primary
aim of our results is to show 1) that nondeterminism at the library
level can exist in real Python programs, 2) that it can be detected
by an implementation of the formalism proposed in this paper, and 3)
that the overhead for such detection, and cost to delta-debug
nondeterministic tests, is not prohibitive.  We believe that the case
studies show that users testing Python libraries, and concerned about
nondeterminism, would find the TSTL tools useful, and would be able to
detect and debug nondeterministic behavior using our approach, a
functionality that we believe to be unique in the automated test generation
literature (and supported in no other tool).  Comparison with a
competing tool is not required to validate the basic idea, since
precise empirical measurements of overheads, or improvements on an
existing approach, are not proposed for evaluation.

Another threat to validity is posed by implementation errors.
TSTL and the mutation tool are tested for bugs by a set of Travis CI
tests, and are (for academic research code) moderately well-used by
external users, making this a lesser concern.  The nondeterminism
detected by TSTL can be demonstrated with standalone tests TSTL
generates that do not depend on any TSTL code.
The source code for the tools and all experimental subjects is
available for inspection and replication in the TSTL repository.