### 2020-01-07 ISSTA/bio.tex

\documentclass{article}

\begin{document}

{\bf Alex Groce:} Alex Groce received his PhD in Computer Science from Carnegie Mellon University in 2005, and worked at NASA's Jet Propulsion Laboratory, as part of the Laboratory for Reliable Software, until 2009, when he joined the faculty at Oregon State University, and was promoted to Associate Professor in 2015.  In 2017, he moved to Northern Arizona University, where he is an associate professor of Computer Science.

{\bf Josie Holmes:}  Josie Holmes graduated from Northern Arizona University in 2004 with a BS in Geography, and received the MSGIS from Pennsylvania State University in 2016.  She is the author of multiple papers in software testing and analysis.

\end{document}