### 2020-01-07 ISSTA/cover.tex

\documentclass{letter}
%\def\today{January 30, 2016}
\usepackage[letterpaper,left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm]{geometry}

\begin{document}
\begin{letter}{}

  \address{Alex Groce\\School of Informatics, Computing \& Cyber Systems\\Northern Arizona University}

  \opening{Dear editor:}

  Attached you will find our submission, ``Does It Always Do That?  Practical Automatic Lightweight Nondeterminism
  and Flaky Test
  Detection and Debugging for Python Libraries.''  This manuscript is original and not in submission elsewhere.  Reviewers knowledgable in delta debugging or random testing systems might be most appropriate for the core content, which concerns the testing mechanics as much as the concepts involved in nondeterminism.

  \closing{Sincerely,\\Alex Groce (agroce@gmail.com)\\Associate
    Professor\\
    School of Informatics, Computing \& Cyber Systems\\Northern Arizona University}

\end{letter}
\end{document}