### 2020-01-07 ISSTA/cover2.tex

\documentclass{letter}
%\def\today{September 10, 2019}
\usepackage[letterpaper,left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm]{geometry}

\begin{document}
\begin{letter}{}

  \address{Alex Groce\\School of Informatics, Computing \& Cyber Systems\\Northern Arizona University}

  \opening{Dear editor:}

  Attached you will find a revision of our submission, ``Does It Always Do That?  Practical Automatic Lightweight Nondeterminism
  and Flaky Test
  Detection and Debugging for Python Libraries.''  We believe this
  revision addresses all reviewer concerns in the two reviews, and
  detail our response in the response to reviewers.

  \closing{Sincerely,\\Alex Groce (agroce@gmail.com)\\Associate
    Professor\\
    School of Informatics, Computing \& Cyber Systems\\Northern
    Arizona University \vspace{0.1in}\\Josie Holmes\\Affiliated Researcher\\
    School of Informatics, Computing \& Cyber Systems\\Northern Arizona University}

\end{letter}
\end{document}