### 2020-01-07 ISSTA/highlights.tex

\documentclass{article}

\begin{document}

\noindent {\bf DOES IT ALWAYS DO THAT? PRACTICAL AUTOMATIC LIGHT- WEIGHT NONDETERMINISM AND FLAKY TEST DETECTION AND DEBUGGING FOR PYTHON LIBRARIES}

\vspace{0.1in}

\noindent {\bf Highlights:}

\vspace{0.1in}

\noindent This article presents a new approach to ``flaky tests'' and nondeterminism.

\vspace{0.1in}

\noindent The article presents a formalism distinguishing horizontal and vertical determinism.

\vspace{0.1in}

\noindent The approach is implemented in an open source Python library testing system.

\vspace{0.1in}

\noindent Failure determinism can detect 6\% more program mutants than a strong specification.

\vspace{0.1in}

\noindent The article presents the first examination of test reduction for probabilistic properties.

\end{document}