### 2020-01-14 ISSTA/abstract.tex
shrink

A critically important, but surprisingly neglected, aspect of system reliability is system predictability.  Many software systems are implemented using mechanisms (unsafe languages, concurrency, caching, stochastic algorithms, environmental dependencies) that can introduce behavioral nondeterminism.  Users of software systems, especially other software systems using a library, often expect that systems will behave deterministically.  Even when it does not impact correctness, nondeterminism must be understood and managed for effective automated testing.  Nondeterministic behavior can result in flaky tests as well as causing problems for test reduction, differential testing, and automated regression test generation.  We show that lightweight techniques, requiring little effort on the part of developers, can extend an existing testing system to allow detection and debugging of nondeterminism.  We show how to make delta-debugging effective for probabilistic faults in general, and that our methods can improve mutation score by 6\% for a strong, full differential test harness for a widely used mock file system.