### 2020-01-16 ISSTA/abstract.tex
big pass

A critically important, but surprisingly neglected, aspect of system reliability is system predictability.  Many software systems are implemented using mechanisms (unsafe languages, concurrency, caching, stochastic algorithms, environmental dependencies) that can introduce unexpected and unwanted behavioral nondeterminism.  Such nondeterministic behavior can result in software bugs and flaky tests as well as causing problems for test reduction, differential testing, and automated regression test generation.  We show that lightweight techniques, requiring little effort on the part of developers, can extend an existing testing system to allow detection and debugging of nondeterminism.  We show how to make delta-debugging effective for probabilistic faults in general, and that our methods can improve mutation score by 6\% for a strong, full differential test harness for a widely used mock file system.