### 2020-01-16 ISSTA/implementation.tex

\section{Implementation}

\begin{comment}
The implementation of the above ideas is conceptually straightforward,
and should be easy to apply in other automated test generation
settings with a natural notion of actions.  We present our
implementation by first sketching a mapping from the formalism to one
particular popular way to represent automatically generated unit
tests, in Section \ref{sec:pools}.  The particular implementations of
horizontal (Section \ref{sec:horizimp}) and vertical (Section
\ref{sec:vertimp}) determinism checks, as well as probabilistic
delta-debugging support (Section \ref{sec:ddimp}) are described after a general
overview of the TSTL testing tool (Section \ref{sec:tstlintro}).
\end{comment}

\subsection{Pools and Visible State}
\label{sec:pools}

While the formal definitions above offer a framework for checking nondeterminism, they leave unspecified the key notions of \emph{state},
\emph{visible state}, and \emph{actions}.  For example, we could
identify the state and visible state of a system as identical with the
full set of memory locations accessed by that system, and actions with
processor instructions.  However, this
definition is highly impractical.  First, it is
extremely inefficient to compare multiple executions for
full-memory-state equivalence at every instruction step.  Second, and
even more importantly, as noted in the introduction, very few programs
are deterministic in this strict sense.  What is a practical level of granularity for (visible) states and actions that
can be efficiently checked, and matches (for the most part) developer
and tester intuitions about what should behave deterministically?

Andrews et al. formalized a value-pool-based model of unit tests that is widely used in
actual testing tools, including Randoop \cite{Pacheco}, showing that
unit tests can be represented in a \emph{canonical} form, where
a test consists of three parts:

\begin{enumerate}
\item A declaration of a set of array variables, referred to as
  \emph{value pools}, e.g., {\tt int [] intVP = new int[3];}.
  \item A set of assignments of constant values to elements of
    primitive type value pools, e.g. {\tt intVP[2] = 1;}.
    \item A part in which all statements are assignments of calls of a
      method to a value pool, with all arguments also
      taken from a pool, e.g. {\tt intVP[2] = fooVP[2].bar(intVP[0],
        intVP[1]);} (referred to as \emph{array-canonical} statements).
    \end{enumerate}

Given such a form for unit tests, there is an obvious mapping to
states and actions in our formalism for nondeterminism:  the state is
the full state of the system, but visible state is restricted to
\emph{the values stored in the value pools}.  Actions are the
array-canonical statements that call code under test to modify the pools.  We can ignore constant assignments and
declarations, since (assuming the language implementation itself is
basically deterministic) these cannot introduce nondeterminism.
Because assertions of correctness and \emph{what a test actually does}
depend on the values in value pools, it seems likely that developers
and testers expect determinism in unit tests, at the granularity of
value pool assignments.

%\subsection{A Brief Introduction to TSTL}
%\label{sec:tstlintro}

We implemented our approach as a modification of the TSTL \cite{NFM15,ISSTA15,tstlsttt}
system, an open-source
language and tool for property-based testing
\cite{Hypothesis,ClaessenH00} of Python code.
%The TSTL {\tt examples} directory
%contains the SUTs needed to replicate our
%experiments.
\begin{comment}
TSTL can also be installed using Python's {\tt pip} tool, by typing
{\tt pip install tstl}; you will need to clone the github repository
to have access to the case studies used in the paper, however.

TSTL has
been used to detect (and usually fix) errors in a number of widely
used Python libraries, the Python implementation itself, the Solidity
compiler and a Solidity static analysis tool, and Mac OS \cite{tstl}.

\begin{figure}
  {\scriptsize
\begin{code}
pool: <int> 5
pool: <l> 5
pool: <s> 5
\vspace{0.1in}
<int> := <[1..20]>
<l> := []
<l> := list(<s>)
<s> := set()
<s> := set(<l>)
\vspace{0.1in}
<l>.append(<int>)
\{ValueError\} <l>.remove(<int>)
<s>.add(<int>)
\{KeyError\} <s>.remove(<int>)
\vspace{0.1in}
property: len(<l>) >= len(set(<l,1>))
property: len(<s>) == len(list(<s,1>))
\end{code}
}
\caption{Simple TSTL harness testing lists and sets}
\label{fig:simple}
\end{figure}
\end{comment}

In TSTL, a test consists of a sequence of
\emph{actions}, where actions execute arbitrary Python code, but are
expected to work by modifying the \emph{state} of a test, which is
stored in a fixed set of pools containing Python objects.  TSTL's
approach essentially follows the pool-based canonical form defined by
Andrews et al. and described above, but relaxes many of its restrictions
(e.g., a constant value can be used in a method call without first
assigning it to a pool). The
TSTL actions defined by a test harness correspond to the set $A$ of
actions in Section \ref{sec:formal}, and transitions between states
result from executing out the Python code in an action; these actions
are, roughly, the array-canonical statements of the value pool
formalism.

\begin{comment}
  Figure
\ref{fig:simple} shows a very simple TSTL harness that tests Python's
built-in {\tt set} and {\tt list} types.  The harness defines the
(visible) state of a test as consisting of three named pools of Python
objects, {\tt int}, {\tt l}, and {\tt s}.  These pools have no
inherent associated type, though a type can be annotated as an
additional constraint that TSTL will check holds during test
execution.  Initially all the pools hold the special Python value {\tt
  None} indicating they have not been initialized.  In terms of our formalism,
$I = \{${\tt [<int> = [None, None, None, None, None]}, {\tt <l> =
  [None, None, None, None, None]}, {\tt <s> = [None, None, None, None,
  None]]}$\}$.  Actions with the {\tt :=} assignment operator
initialize pools.  For example, the line of code {\tt l := []} defines a set of 5
actions, each of which initializes a different member of the list pool
{\tt l} to an empty list --- {\tt l0 = []} ... {\tt l4 = []}.  Actions
without a {\tt :=} require that pools used in them already be
initialized.  TSTL provides constructs like {\tt <[1..20]>} for
choosing values in a range for use in a test, and allows a user to
specify which exceptions an action may raise without causing the test
to fail (as in {\tt \{KeyError\} <s>.remove(<int>)}).

Compiling this harness using
the TSTL compiler produces a Python module that allows test
generation.  If we use the random testing tool provided by TSTL, {\tt
  tstl\_rt}, it will randomly generate valid sequences of actions, and
report no problems.  If we change the first property to read {\tt
  property: len(<l>) == len(set(<l,1>))}, however, TSTL will report a
failing test, shown in Figure \ref{fig:setlistfail}.

\begin{figure}
{\scriptsize
\begin{code}
int0 = 1                                                                 \# STEP 0
ACTION: int0 = 1 
int0 = None : <type 'NoneType'>
=> int0 = 1 : <type 'int'>
==================================================
l0 = []                                                                  \# STEP 1
ACTION: l0 = [] 
l0 = None : <type 'NoneType'>
=> l0 = [] : <type 'list'>
==================================================
l0.append(int0)                                                          \# STEP 2
ACTION: l0.append(int0) 
int0 = 1 : <type 'int'>
l0 = [] : <type 'list'>
=> l0 = [1] : <type 'list'>
==================================================
l0.append(int0)                                                          \# STEP 3
ACTION: l0.append(int0) 
int0 = 1 : <type 'int'>
l0 = [1] : <type 'list'>
=> l0 = [1, 1] : <type 'list'>
==================================================
ERROR: (<type 'exceptions.AssertionError'>,
  AssertionError(), <traceback object at 0x1032b83f8>)
TRACEBACK:
  File "/Users/adg326/scratch/pyexample/sut.py",
  line 13695, in check
    assert len(self.p\_l[0]) == len(set(self.p\_l[0]))
\end{code}
}
\caption{A failing test (showing that converting a list to a set may
  not preserve the size of the object).}
\label{fig:setlistfail}
\end{figure}

\end{comment}

\subsection{Nondeterminism Detection}
%\label{sec:horizimp}

Because TSTL supports differential testing \cite{tstlsttt}, horizontal
nondeterminism detection can technically be implemented simply by
declaring a system to be its own reference, using TSTL's notation for
differential testing.  However, such an approach requires considerable effort
on the part of the user to express which values are checked for
equivalence, and does not (without a great deal of effort) support injecting timing differences,
or re-executing in a new process, in checking for nondeterminism.

We therefore instead made horizontal nondeterminism detection a first-class
property in TSTL, using TSTL's existing notation for marking some
types of values as \emph{opaque} (not usefully compared for equality),
used in automatic abstraction-based testing.  Recall that in TSTL, a
test is (essentially) a sequence of assignments to pool values,
and method/function calls using pool values (including method calls on
pool objects).  In TSTL, \emph{visible
value determinism} is, as proposed above, based on comparing the values of \emph{all pool
variables} after each test action (pool values other than the one
assigned to must also be compared, since a common source of
nondeterminism is when a call results in a change to a value passed as
a parameter, or changes a shared reference held by two values).  By
default $V$ as defined in Section \ref{sec:formal} is just the set of all possible pool values, and $v$
simply extracts the pool values from $S$.
TSTL allows restriction of which state is visible by removing pools marked as {\tt OPAQUE} in the
TSTL language test harness (and values that do not support equality
checking) from $v(S)$.  \emph{Final State Determinism} simply performs the same
comparison, but only on the final values of all pool variables (or a
set of designated pools) after a test has
finished executing.

\begin{comment}
\begin{table*}
\centering
{\scriptsize

\caption{TSTL Method Calls for Nondeterminism Detection}
\label{tab:methods}
\begin{tabular}{l|p{8cm}}
%\hline
Name & Purpose \\
\hline
\hline
{\tt nondeterministic} & returns True if test exhibits final state nondeterminism in $k$
  tries with delay $d$, optionally only over pools in $\mathcal{P}$\\
\hline
{\tt stepNondeterministic} & as above, except checks all visible values
                       (compares state at every step)\\
\hline
{\tt processNondeterministic} & as above, except runs test provided in
                          subprocess and compares non-opaque output
                          values \\
\hline
{\tt P} & returns probability that a predicate holds for a test (with
          optional number of samples to use in estimate) \\
\hline
{\tt forceP} & wraps a predicate for test reduction, checking that it
               has desired probability of holding, using
         sample size and replications parameters \\
\hline
\hline

%\hline
\end{tabular}
}
\end{table*}

\begin{table*}
\centering
{\scriptsize
\caption{TSTL Command Line Options for
  Nondeterminism Detection}
\label{tab:options}
{\scriptsize
\begin{tabular}{l|l|p{6cm}}
%\hline
Name &  Tool(s) & Purpose \\
\hline
\hline
{\tt --checkDeterminism} & {\tt rt}, {\tt reduce} & Checks tests for deterministic
                                  behavior (visible value in rt, final
  value in reduce)\\
\hline
{\tt --checkProcessDeterminism} & {\tt rt}, {\tt reduce} & Checks determinism using more 
                                         expensive process-changed 
                                         execution \\
\hline
{\tt --determinismTries INT} & {\tt rt}, {\tt reduce} & Number of times to repeat test 
                                  checking for nondeterministic
                                  behavior \\
\hline 
{\tt --determinismDelay FLOAT} & {\tt rt}, {\tt reduce} & Delay between steps when 
                                  re-executing to check for
                                              determinism \\
\hline 
{\tt --checkFailureDeterminism} & tstl compiler & Produce a harness
                                                  that checks
                                                  determinism on all
                                                  failing actions \\
\hline 
{\tt --probability P} & {\tt reduce} & force property with respect to which 
                               test is being reduce to hold with 
                                 probability $P$ \\
\hline 
{\tt --samples S} & {\tt reduce} & use $S$ samples when checking predicate
                             probability\\
\hline 
{\tt --replications R} & {\tt reduce} & use $R$ replications when checking predicate probability\\
\hline
\hline

%\hline
\end{tabular}
}
}
\end{table*}

 Tables \ref{tab:methods}
and \ref{tab:options} show, respectively, the primary additions to the
TSTL Python API and the TSTL command line tools; in addition to
horizontal nondeterminism checking, these extensions also support failure
nondeterminism and probabilistic delta-debugging.  The TSTL test
generation tool is called {\tt rt}, since the default mode is a fast
pure Random Tester, and {\tt reduce} is TSTL's tool for
delta-debugging and test
normalization (more aggressive reduction that also tends to ease
debugging and test triage by making tests failing due to the same
fault more similar \cite{onetest}).  With these additions,
developers of TSTL-based testing tools, or developers using TSTL tools
to test code can easily add nondeterminism to the set of properties
checked.  It is even possible to write a TSTL harness that
automatically checks itself for nondeterminism, even if the random
tester is not run with {\tt --checkDeterminism}, simply by adding:

{\scriptsize
\begin{code}
property: not self.nondeterministic(self.test())
\end{code}
}

\noindent to the harness; in fact, by removing the {\tt not}, a test harness can
even specify that behavior should not be deterministic, a property of
possible use in some security-related libraries.

\end{comment}

We provide support for the real-world sources of
nondeterminism described in Section \ref{sec:sources}, via
control over timing and/or execution in a new process.  Changing the timing of
steps is also effective in exposing concurrency-related
nondeterminism, since it often changes the
interleaving of threads.

Because TSTL has an interface to AFL \cite{aflfuzz}, we can use
AFL's sophisticated heuristics to perform very thorough, week-long
checks for nondeterminism.  Moreover, since
there is substantial overhead in nondeterminism detection, AFL can be
used to predict whether tested Python code is a good candidate for
checking for horizontal nondeterminism; if the
AFL \emph{stability} statistic (see the AFL documentation for details
\cite{aflfuzz}) is lower than 100\%, it may indicate horizontal
nondeterminism.
Because of the idiosyncrasies of AFL instrumentation and process
behavior, this is not always a reliable guide, but it is a very low
cost indicator produced as a by-product of fuzzing for other problems.

%\subsection{Vertical Nondeterminism Detection}
%\label{sec:vertimp}

First-class vertical nondeterminism checks are currently
limited to failures.  
%{\tt --checkFailureDeterminism} causes the TSTL
%compiler to emit a harness where every action that causes an \emph{expected}
%exception to be raised is repeated to ensure the same exception is
%still raised.
In terms of the formalism, if execution of action
$\alpha_i$ raises an exception that does not, itself, indicate a test failure, then 
    $F_i$ is true, and TSTL repeats $\alpha_i$ in order to ensure that
    $F_{i+1}$ is also true.
    %\end{enumerate}

%When a failure is indicated by some means other than raising an
%exception, TSTL users must express this by explicitly writing
%a wrapper that raises an expected exception on failure.

\begin{comment}
\subsection{Tools for Nondeterministic Delta-Debugging}
\label{sec:ddimp}

Unlike the other additions, explicitly designed for nondeterminism
detection, the {\tt P} query, {\tt forceP} wrapper, and {\tt --probability}, {\tt
  --samples}, and {\tt --replications} reducer options are general
tools for use with nondeterministic predicates in delta-debugging.  Of
course, nondeterminism and flakiness are the most common probabilistic
test properties we are aware of, but in theory these tools can be used
for other probabilistic debugging problems (e.g., if the property of
interest is that a value is above a certain threshold in 90\% of test
executions, these TSTL additions can be used to find a small test where this property
fails to hold).  In addition to the basic probabilistic reduction
issues discussed above, actual implementation of test reduction for
nondeterminism introduces some subtle issues.  By default, for
example, TSTL automatically removed non-enabled actions from tests
during reduction, because these are never actually executed.  However,
in timing based nondeterminism, such steps may be essential to
causing the nondeterminism to appear (we had to turn off this feature
for the {\tt redis-py} example discussed below).

\subsection{Key Challenges}
The key challenge in all of these implementations was choosing
appropriate default values.  In particular, how many times to retry
tests (horizontal determinism) or actions (vertical/failure determinism), what delays to use, and numbers of replications
required in nondeterministic delta-debugging.  Because our confidence
that these defaults are ideal is low, we made it easy to change the
values used, as shown by the presence of command line options and
optional function parameters in Tables \ref{tab:methods} and
\ref{tab:options}.
\end{comment}

\begin{comment}
\section{A Simple Example}

\begin{figure}
{\scriptsize
\begin{code}
@from storage import *
\vspace{0.1in}
init: init()
pool: <int> 5
\vspace{0.1in}
<int> := <[-10..10]>
<int> := getNewVal()
\{KeyError,ValueError\} store(<int>)
\vspace{0.1in}
property: len(contents()) == len(set(contents()))
\end{code}
}
\caption{TSTL harness for Python library with (independent) vertical and
  horizonal nondeterminisms}
\label{fig:simpleexample}
\end{figure}

\begin{figure}[t]
  {\scriptsize
\begin{code}
import random                                   def getNewVal():
storage = []                                            global storage
def contents():                                        if len(storage) >= 1:
    global storage
    return storage
def init():
    global storage
    storage = []
def store(n):
    global storage
    if n in storage:
       raise KeyError
    storage.append(n)
    if n < 0:
       raise ValueError

def getNewVal():
    global storage
    if len(storage) >= 1:
        bot = min(storage)
        top = max(storage)
    else:
        bot = -10
        top = 10
    v = random.randint(bot-1,top+1)
    while v in storage:
        v = random.randint(bot-1,top+1)
    return v
  \end{code}
  }
\caption{Code for Python library with (independent) vertical and
  horizonal nondeterminisms}
\label{fig:simpleexamplecode}
\end{figure}

Figures \ref{fig:simpleexample} and \ref{fig:simpleexamplecode} show a self-contained example of a
Python library and test harness with both vertical and horizontal
nondeterminism.  The example is not meant to be realistic, but to
better concretely explain the detection and debugging of
nondeterminism.  A few additional TSTL features need to be described.
First, lines beginning with an {\tt @} indicate raw Python code, here
used to import the module to be tested.  Second, {\tt init:} indicates
code that is to be executed at the beginning of each test, in
order to restore the system under test to its initial state.

\subsection{Horizontal Nondeterminism}

If we compile and test this code without using any of the
nondeterminism-detection functionality, TSTL will not report any
failing tests.  The property (avoidance of duplicate entries in
storage) holds:

{\scriptsize
\begin{code}
 \$ tstl storage.tstl
 ...
 \$ tstl\_rt
 Random testing using config=Config...
 storage: TEST \#1396 STEP \#0 (0:00:30.000188) Mon Feb
 25 11:13:18 2019 [ 15 stmts 22 branches ] (no cov+ for
 1391 tests) 139500 TOTAL ACTIONS (4649.97/s) (test
 0.0/s) 2 cov+ tests
\end{code}
}

TSTL reports information on the testing in a status line, here showing
that it has performed 1,396 tests, with no failures, and two tests
that increased code coverage.  TSTL can be run in a mode that stores
minimized versions of the tests that obtained new coverage, for use
as, e.g., regressions:

{\scriptsize
\begin{code}
 \$ tstl storage.tstl
 ...
 \$ tstl\_rt --quickTests
\end{code}
}

\begin{figure}
 {\scriptsize
{\bf quick.0.test:}
\begin{code}
self.p\_int[0] = getNewVal() 
store(self.p\_int[0]) 
store(self.p\_int[0]) 
self.p\_int[0] = -10 
store(self.p\_int[0]) 
\end{code}
\vspace{0.1in}
{\bf quick.1.test:}
\begin{code}
self.p\_int[0] = -8
store(self.p\_int[0]) 
self.p\_int[0] = getNewVal() 
\end{code}
}
\caption{Coverage maximizing tests for Python library with (independent) vertical and
  horizonal nondeterminisms}
\label{fig:simpleexampletests}
\end{figure}

This will result in generation of two\footnote{In some rare cases,
  TSTL may generate one test that achieves complete code coverage, but
  this will only happen rarely here.} files, {\tt quick.0.test} and
{\tt quick.1.test}.  Sometimes, TSTL will fail at this point with an internal error
({\tt REDUCED TEST DOES NOT PRESERVE COVERAGE}), due to the attempt to reduce one of the coverage-obtaining
tests losing coverage, which is a consequence of horizontal
nondeterminism.  However, TSTL will often produce
test files such as those shown in Figure
\ref{fig:simpleexampletests}.  TSTL lets us convert these tests, which
are in TSTL's internal format, into standalone Python regression tests
that check the properties in the test harness and ensure that the
behavior (values of pools) are the same as in the recorded run:

{\scriptsize
\begin{code}
 \$ tstl\_standalone quick.0.test quick0.py --regression
 ...
 \$ tstl\_standalone quick.1.test quick1.py --regression
\end{code}
}

However, if we execute these Python tests, they will be very flaky;
sometimes they will pass, but {\tt quick0.py} will almost always fail, and
{\tt quick1.py} will fail about two thirds of the time.  The reason,
of course, is a trivial case of horizontal nondeterminism, due to the
use of the {\tt random.randint} function in {\tt getNewValue}.  The
regression test will record whatever value {\tt getNewValue} returned
the when the TSTL test file was executed in order to produce the
Python standalone test.  The value in future executions will seldom
match this value.  In this case, removing the nondeterminism is easy
(simply add a fixed random seed to the {\tt init:} line in the
harness, e.g., {\tt init: init(); random.seed(0)}).  However, in
real-world testing, the source of the nondeterminism, may not be
obvious, and only some regression tests may (infrequently) fail.
However, if we ask the random tester to check for nondeterminism, it
quickly shows us what the problem is:

{\scriptsize
\begin{code}
 \$ tstl\_rt --checkDeterminism
 ...
 int2 = getNewVal()              \# STEP 50
 MISMATCH IN REPLAY VALUE:
    int2 : 7 VS. -5
 TEST WAS NOT DETERMINISTIC!  WRITING FAILURE AS
 ndfail.test
\end{code}
}

We can use TSTL's test reduction tool to produce a small test showing
the problem:

{\scriptsize
\begin{code}
 \$ tstl\_reduce ndfail.test hnd.test --checkDeterminism
 ...
 NEW LENGTH 1
 int0 = getNewVal()              \# STEP 0

 TEST WRITTEN TO hnd.test
\end{code}
}

Nothing is required to produce nondeterminism other than calling {\tt
  getNewVal}, and this test is more frequently nondeterministic than
the original, since it calls {\tt getNewVal} without any stored data
to restrict the range of random values.

\subsection{Vertical Nondeterminism}

The {\tt storage.py} program also contains an (again, unrealistically
simple) example of \emph{vertical} nondeterminism.  The {\tt store}
method is supposed to 1) allow only a single copy of a value to be in
the storage and 2) to disallow negative values.  Unfortunately, the
check for negatives is \emph{after} the value is stored, which means
that a second attempt to store the same negative value will result in
a {\tt KeyError}, not a {\tt ValueError}.  TSTL can detect this
problem if we compile the harness with {\tt
  --checkFailureDeterminism}, which produces a version of the testing
interface that ensures failing actions (ones that raise an expected
exception) always fail again if performed again without any
intervening actions, and always fail with the same exception.  Note
that the check for the same exception is a slight extension of our formal
definition of failure determinism, recording the last observed
exception as an additional visible state element.

{\scriptsize
\begin{code}
 \$ tstl storage.tstl --checkFailureDeterminsim
 ...
 \$ tstl\_rt
 ...
 FINAL VERSION OF TEST, WITH LOGGED REPLAY:
 int0 = -4                                   \# STEP 0
 ACTION: int0 = -4 
 int0 = None : <type 'NoneType'>
 => int0 = -4 : <type 'int'>
 ==================================================
 store(int0)                                 \# STEP 1
 ACTION: store(int0) 
 int0 = -4 : <type 'int'>
 RAISED EXPECTED EXCEPTION:
   <type 'exceptions.ValueError'> 
 RAISED EXCEPTION:
   <type 'exceptions.KeyError'>  ON RETRY
   ERROR: (<type 'exceptions.AssertionError'>,
   AssertionError('Failure with different exception.',),
 <traceback object at 0x105e9edd0>)
\end{code}
}

With failure nondeterminism, we don't have to use TSTL's specialized reduction
tool, since the failure does not require multiple replays of a test.
The random tester itself can properly reduce the test case, since it
is just a ``normal'' failure, exhibited by a single trace.

\end{comment}