### 2020-07-15 ISSTA/relatedwork.tex

\section{Related Work}

The general topic of nondeterminism in software engineering has long been considered important  \cite{Unknowns,UncertaintyPrinciple}. 
Gao et al. examined the question of how to make tests repeatable \cite{Gao:2015:MSU:2818754.2818764}, in the context of system-wide user interaction tests, focusing on systemic factors.  
Shi et. al \cite{DetermImp} examined code that wrongly enforces a deterministic implementation.  %E.g., they detect instances when  iteration through items of a set is explicitly not guaranteed to provide any particular order of items, but code depends on the order. Determinism is also sometimes used as a property in domain-specific testing, e.g. for shader compilers \cite{shader}.

The problem of test nondeterminism is closely related to the issue of flaky tests \cite{miccoflaky, luo2014empirical,palomba2017does,listfieldtestanalysis}.  How to handle flaky tests in practice is a major issue in Google-scale continuous testing \cite{memon2017taming}.
Previous work on flaky tests has either focused on test inter-dependence \cite{LamZE2015}, or large-scale empirical examination \cite{luo2014empirical,palomba2017does}.  
Bell et al. proposed DeFlaker \cite{bell2018d}, which makes flaky tests much easier to detect by relying on the observation that if a test fails, and does not cover any changed code then the test is likely flaky.   iDFlakies \cite{idflakies} is a framework and dataset for flaky tests, but again focuses on whole-tests, and detecting flakiness by actually observing it.

We introduce the first variation of
delta-debugging that properly handles probabilistic reduction criteria, in line with Harman and O'Hearn's proposal to simply accept that ``All Tests Are Flaky'' \cite{StartupstoScaleups}, and work with probabilistically failing tests..
Vertical/failure determinism is also a less-studied concept, and to our knowledge our formulation is novel. The kinds of errors that are exposed, however, are not new, e.g., faults related to the propagation of error conditions \cite{FileProp}.

%TSTL's approach to test generation, and our instantiation of the formal definitions of nondeterminism are based on a formulation of unit tests using \emph{pools} of values \cite{AndrewsTR}, which provides a practical solution to the problem of defining the visible state to be compared when checking for nondeterminism.